---
output: github_document
---

<!-- README.md is generated from README.Rmd. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

if (!dir.exists(file.path("inst", "extdata")))
  dir.create(file.path("inst", "extdata"), recursive = TRUE)
```

```{r, echo = FALSE}
library(soilDB)
options(soilDB.warn.aliases = FALSE)
```

# {soilDBdata}

<!-- badges: start -->
<!-- badges: end -->

The goal of {soilDBdata} is to provide a standard way of building test data sets for the soilDB R package. Most of these test datasets are in the form of SQLite 'snapshots' of NASIS tables.

These data sets are regularly updated to support unit testing of [NASIS functionality in the {soilDB} package](https://github.com/ncss-tech/soilDB/blob/master/tests/testthat/test-soilDBdata.R). A few data sets housed in this repository are also used for training courses provided by the Soil and Plant Science Division such as [Statistics for Soil Survey](https://ncss-tech.github.io/stats_for_soil_survey/book/index.html).

Access to the NASIS transactional database requires eAuth and a NASIS user account. NASIS data are intended for internal use by NRCS staff and authorized cooperators. NASIS-related functions in {soilDB} provide a standard read-only connection to a local NASIS instance for users that have such access, or alternately can be supplied with the optional `dsn` argument. The `dsn` argument can be a _DBIConnection_ to an MSSQL, SQLite or PostgreSQL database with the NASIS schema (for instance), or a path to a SQLite database file. 
 
## Installation

You can install the released version of {soilDBdata} from GitHub with:

``` r
remotes::install_github("brownag/soilDBdata")
```

## Example 1 `soilDB::fetchNASIS('pedons')`

This is an example that shows you how to create a snapshot of the tables in your local database for running `fetchNASIS()` without a local NASIS database. 

The sample data for testing the `from="pedons"` functionality are queried into the local database, accepted, and queried into and selected set using SSRO_Northwest query `PedonPC_Plus_DataDump_select` targeting Site and Pedon objects. User Site ID matches `"2015MT663%"`, NASIS site name matches `"%"` and NASIS group name matches `"NW-MIS Point Data"`. 

A total of `134` Site and Pedon objects should be found, and `115` of those profiles are retained in `fetchNASIS()` with `rmHzErrors = TRUE` because `19` profiles do not have horizons. These profiles represented with a single `NA` depth horizon when `fill = TRUE`.

<!-- The following groups of tables are used by `fetchNASIS()` and related Site/Pedon-level queries in {soilDB}: -->
<!-- ```{r echo=FALSE} -->
<!-- # NASIS Table -->
<!-- knitr::kable(data.frame(soilDB::get_NASIS_table_name_by_purpose( -->
<!--   purpose = c("site", "pedon", "transect", "metadata", "lookup") -->
<!-- )), col.names = "Table") -->
<!-- ``` -->

Create a "static" snapshot for `soilDB::fetchNASIS('pedons')` from your local NASIS database.

```{r example1}
library(soilDBdata)

con <- soilDB::NASIS()
rebuild <- FALSE
dsn <- "inst/extdata/fetchNASIS_pedons.sqlite"

if (rebuild && inherits(con, 'DBIConnection')) {
  res <- create_fetchNASIS_pedons(output_path = dsn)
  DBI::dbDisconnect(con)
}

if (file.exists(dsn)) {
  
  f <- try(soilDB::fetchNASIS(
    from = "pedons",
    dsn = dsn,
    SS = TRUE,
    rmHzErrors = TRUE
  ))
  
  f2 <- try(soilDB::fetchNASIS(
    from = "pedons",
    dsn = dsn, 
    SS = FALSE, 
    fill = TRUE
  ))
  
}

if (!is.null(f))
  show(f)

if (!is.null(f2))
  show(f2)
```

## Example 2 `soilDB::fetchNASIS('components')`

Create a "static" snapshot for `soilDB::fetchNASIS('components')` from your local NASIS database.

The sample data for testing the `from="components"` functionality are queried into the local database, accepted, and queried into selected set using SSRO_Southwest query `Whole multiple legends by area sym != addit. and DMU = yes +` targeting Area, Legend, Mapunit and Data Mapunit objects, where Area Symbol matches `"MT663"`. 

A total of `1` Legend, `82` Mapunit and `86` Data Mapunit objects should be downloaded into local database, with `63` (non-additional) Mapunit and `63` Data Mapunit objects loaded into selected set. `fetchNASIS(from="components")` will return `191` component profiles with `fill=FALSE` and `546` with `fill=TRUE`.

```{r example2}
library(soilDBdata)

con <- soilDB::NASIS()
rebuild <- FALSE
dsn <- "inst/extdata/fetchNASIS_components.sqlite"

if (rebuild & inherits(con, 'DBIConnection')) {
  res <- create_fetchNASIS_components(output_path = dsn)
  DBI::dbDisconnect(con)
}

if (file.exists(dsn)) {
  f <- try(soilDB::fetchNASIS(
    from = "components",
    dsn = dsn,
    SS = TRUE,
    rmHzErrors = TRUE
  ))
  
  f2 <-  try(soilDB::fetchNASIS(
    from = "components",
    dsn = dsn,
    SS = FALSE,
    fill = TRUE
  ))
}

if (!is.null(f))
  show(f)

if (!is.null(f2))
  show(f2)
```

## Example 3 `soilDB::fetchVegdata()`

Create a "static" snapshot for `soilDB::fetchVegdata()` from your local NASIS database.

The sample data for testing the `fetchVegdata()` functionality are queried using two queries to the national database. First run `POINT - Pedon/Site by User Site ID` with target table "Pedon" to download "Site", "Pedon", and "Transect" objects where User Site ID matches`"2010WY629%"`.. Then run `VEG - Veg Plots by User Site ID and Vegplot ID` with target table "Vegetation Plot" to get "Site" and "Vegetation Plot" objects where User Site ID and Vegetation Plot ID matches `"2010WY629%"`.

The first query will result in `26` Site and `26` Pedon objects. Of the `26` pedons, `11` have vegetation plots linked to the same site observation. The second query will result in `137` Site, and `138` Vegetation Plot objects. One Site Observation has two Vegetation Plot objects linked to it. All `138` Vegetation Plots have `assocuserpedonid` populated, but most of the pedons do not exist in NASIS. 
The two queries result in `152` unique sites matching `"2010WY629%"`, of which `127` lack Pedons and `15` lack Vegetation Plots.

```{r example3}
library(soilDBdata)

con <- soilDB::NASIS()
rebuild <- FALSE
dsn <- "inst/extdata/fetchVegdata.sqlite"

if (rebuild && inherits(con, 'DBIConnection')) {
  res <- create_fetchVegdata(output_path = dsn)
  DBI::dbDisconnect(con)
}

if (file.exists(dsn)) {
  f <- try(soilDB::fetchVegdata(
    dsn = dsn, 
    SS = TRUE
  ))

  f2 <- try(soilDB::fetchVegdata(
    dsn = dsn, 
    SS = FALSE,
    include_pedon = "assocuserpedonid"
  ))
  
  f3 <- try(soilDB::fetchVegdata(
    dsn = dsn, 
    SS = FALSE,
    include_pedon = FALSE
  ))
}

if (!is.null(f))
  str(f, max.level = 1)

if (!is.null(f2))
  str(f2, max.level = 1)

if (!is.null(f3))
  str(f3, max.level = 1)
```
